---
title: '[STAT 4400] HW-5'
author: "Michael Ghattas"
date: "3/29/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,out.width="0.9\\linewidth",dev="png",fig.align  = 'center')
pacman::p_load(
ggplot2,
knitr,
arm,
data.table,
foreign,
gridExtra,
car,
stringr,
rstan,
rstanarm,
zoo
)
library(knitr)
```

# Problem - 1

```{r}
library(lme4)
library(ggplot2)

df <- read.table(file = "/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/apt.txt", header = TRUE)
head(df)
```

### (a)

```{r}
m1 <- lmer(y ~ dist + (0 + dist | race), data = df)
summary(m1)
```

### (b)

```{r}
m2 <- lmer(y ~ dist + (1 + dist | race), data = df)
summary(m2)
```

# Problem - 2

```{r}
library(reshape)
library("stringr")
library(zoo)

filename <- "/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/olympics1932.txt"
olympics1932_na <- read.fwf(filename, widths = c(2, 14, 9, 9, 9, 9, 9, 9, 9), skip = 21, header = FALSE)
colnames(olympics1932_na)<- c("pair", "criterion", "judge_1",  "judge_2",  "judge_3", "judge_4",  "judge_5" , "judge_6",  "judge_7")
olympics1932 <- na.locf(olympics1932_na)
olympics1932$criterion <- str_trim(olympics1932_na$criterion)
```

### (a)

```{r}
arr_olym <- melt(data = olympics1932, id.vars = c("pair", "criterion"), measure.vars=c(colnames(olympics1932)[3:9]))
arr_olym
```

### (b)

```{r}
olym_984 <- arr_olym[order(arr_olym$variable), ]
olym_984 <- olym_984[c("criterion", "value", "pair", "variable")]

olym_984
```

### (c)

```{r}
olym_984$SameCountry <- ifelse(olym_984[, 3] == " 1"&olym_984[, 4] == "judge_5", 1, ifelse(olym_984[, 3] == " 2"&olym_984[, 4] == "judge_7", 1, 
  ifelse(olym_984[, 3] == " 3"&olym_984[, 4] == "judge_1", 1, ifelse(olym_984[, 3] == " 4"&olym_984[, 4] == "judge_1", 1, 
  ifelse(olym_984[, 3] == " 7"&olym_984[, 4] == "judge_7", 1, 0)))))

olym_984
```

# Problem - 3

```{r}
library(tidyverse)
library(ggvis)
library(data.table)

hiv.data.raw <- fread("/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/allvar.csv")

invisible(hiv.data.raw[,ok := !is.na(CD4PCT) ])
hiv.data<-hiv.data.raw[ok==TRUE]
invisible(hiv.data[,y :=sqrt (CD4PCT)])
 # kid's age (yrs) at the beginning of the study
invisible(hiv.data[,age.baseline := baseage ]  ) 
# kids age (yrs) at the time of measurement
invisible(hiv.data[,age.measurement := visage ] ) 
invisible(hiv.data[,time := visage - baseage ] )
setnames(hiv.data,"treatmnt","treatment") 
hiv.data<-hiv.data[complete.cases(hiv.data[,list(y,time,age.baseline,treatment)])]
```

### (a)

```{r}
ggplot(hiv.data) + geom_point(aes(x = time, y = y)) + xlab("Time") + ylab("Outcome (the CD4 percentage, on the square root scale)")
```

### (b)

```{r}
r_np <- lm(y ~ time + factor(newpid) - 1, data = hiv.data)
summary(r_np)

ggplot(aes(x = time, y = y), data = hiv.data) + geom_point() + geom_abline(intercept = coef(r_np)[1], slope = coef(r_np)[2:length(coef(r_np))]) + 
  xlab("Time") + ylab("Outcome (the CD4 percentage, on the square root scale)")
```

### (c)

```{r}
r1 <- lm(y ~ time + factor(newpid) - 1, data = hiv.data)
summary(r1)

child <- hiv.data %>% select(newpid, age.baseline, treatment)
child <- unique(child)

r1.coef <- data.frame(child, r1$coefficients[2:length(r1$coefficients)])
colnames(r1.coef) <- c("newpid", "age.baseline", "treatment","coef.id")
rownames(r1.coef) <- 1:250
r1_coef.id <- lm(coef.id ~ age.baseline + factor(treatment), data = r1.coef)

summary(r1_coef.id)
```

# Problem - 4

### (a)

```{r}
M0 <- lmer (y ~ time + (1 | newpid), data = hiv.data)
display(M0)

M0.coef <- data.frame(unique(hiv.data$newpid),coef(M0)$newpid)
colnames(M0.coef) <- c("newpid", "intercept", "time")
head(coef(M0)$newpid)
```

The coefficient for time is ~(-0.3661), constant across the children. Thus if the time increases by 1 unit, then the CD4 percentage on the square root scale will decrease by ~(0.3661) units.

### (b)

```{r}
M1 <- lmer (y ~ time + factor(treatment) + age.baseline + (1 | newpid), data = hiv.data)
display(M1)

head(coef(M1)$newpid)
```

1. The coefficients for time, treatment and age.baseline are all constant across the children
2. The coefficient for time is ~(-0.36216), thus if the time increases by 1 unit, then CD4 percentage on the square root scale will decrease by ~(0.36216) units
3. The coefficient for treatment is ~(0.18008), thus the CD4 percentage on the square root scale for children under treatment 2 is ~(0.18008) more than treatment 1
4. The coefficient for time is ~(-0.11945), thus if the age.baseline increases by 1 unit, CD4 percentage on the square root scale will decrease by ~(0.11945) units

### (c)

```{r}
data_plot <- as.data.frame(cbind(unlist(ranef(M0)), unlist(ranef(M1))))
colnames(data_plot) <- c("M0", "M1")
ggplot(data = data_plot,aes(x = M0,y = M1)) + geom_point() + geom_smooth() +  xlab("Random effcts of M0") +  ylab("Random effects of M1")

display(M0)
display(M1)
```

### (d)

```{r}
M2<-lmer(hiv.data$y ~ hiv.data$time + (1 + hiv.data$time | hiv.data$newpid))
M3<-lmer(hiv.data$y ~ factor(hiv.data$time) + (1 | hiv.data$newpid))

data_plot2_inter <- as.data.frame(cbind(unlist(ranef(M2))[1:250], unlist(ranef(M3))[1:250]))
colnames(data_plot2_inter) <- c("M2", "M3")
ggplot(data = data_plot2_inter, aes(x = M2,y = M3)) + geom_point() + geom_smooth() + 
  xlab("Random effcts of M2 intercepts") + ylab("Random effects of M3 intercepts")

display(M2)
display(M3)
```

# Problem - 5

```{r}
one <- lmer(formula = hiv.data$y ~ hiv.data$time + hiv.data$age.baseline + hiv.data$treatment + (1 | hiv.data$newpid))

summary(one)
plot(one)
```

##### Original formula of the multilevel model:

$$y=\beta_{0[j]i}+X_{i1}*\beta_{1[j]i}+X_{i2}*\beta_{2[j]i}+X_{i3}*\beta_{3[j]i}+\epsilon_i \, $$

$$y=\alpha_{j[i]}+\beta_1X_{i1}+\beta_2X_{i2}+\beta_3X_{i3}+\epsilon_i\, $$

$$\alpha_j\sim\ {N}(\mu_i,\sigma_i^2)\, $$

$$X1=time,\ X2=age.baseline,\ X3=treatment$$.

##### Method1: Allowing regression coefficients to vary accross groups

$$y=4.91+X_{i1}*(-0.36)+X_{i2}*(-0.12)+X_{i3}*0.18+0.77 \ ,\ for\ i=1,...,n_{250} $$

$$\alpha_j\sim\ \mathrm{N}(0,1.37^2)\\$$

##### Method2: Combining separate local regressions

$$y\sim\ N(4.91+X_{i1}*(-0.36)+X_{i2}*(-0.12)+X_{i3}*0.18,\ 0.77^2)\ ,\ for\ i=1,...,n_{250}$$

$$\alpha_j\sim\ \mathrm{N}(random\ intercept,1.37^2)\\$$

##### Method3: Modeling the coefficients of a large regression model

$$y_i \sim\ N(4.91+X_{i1}*(-0.36)+X_{i2}*(-0.12)+X_{i3}*0.18,\ 0.77^2)$$

$$\beta_j\sim\ N(0,1.37^2)$$

##### Method4: Regression with multiple error terms

$$y_i \sim\ N(4.91+X_{i1}*(-0.36)+X_{i2}*(-0.12)+X_{i3}*0.18+1.37^2,\ 0.77^2)$$

##### Method5: Large regression with correlated errors

$$y_i \sim\ N(4.91+X_{i1}*(-0.36)+X_{i2}*(-0.12)+X_{i3}*0.18,\ 1.37^2+0.77^2)$$