---
title: '[STAT 4400]  HW-3'
author: "Michael Ghattas"
date: "2/15/2022"
output:
  word_document: default
  html_document: default

reference: https://github.com/IamGianluca
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

```{r}
require(AER)
require(arm)
require(foreign)
require(ggplot2)

df <- read.dta("/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/risky_behaviors.dta", convert.factors = TRUE)
df$fupacts <- round(df$fupacts)
df$couples <- factor(df$couples)
df$women_alone <- factor(df$women_alone)

summary(df)
```


### (a)

```{r}
poi.reg <- glm(fupacts ~ women_alone, family=poisson, data = df)
summary(poi.reg)
```

The model is a poor fit, even with the woman_alone factor having a statistical significance.


### (b)

```{r}
df$c.bupacts <- (df$bupacts - mean(df$bupacts)) / (2 * sd(df$bupacts))
poi.reg.ext <- glm(fupacts ~ women_alone + sex + c.bupacts + couples + bs_hiv, family = poisson, data = df)
summary(poi.reg.ext)

par(mfrow = c(1,2))
plot(predict(poi.reg.ext), residuals(poi.reg.ext), xlab = "Predicted", ylab = "Residuals")
abline(a = 0, b = 0, lty = 3)
plot(predict(poi.reg.ext), rstandard(poi.reg.ext), xlab = "Predicted", ylab = "Standardized Residuals")
abline(a = 1.96, b = 0, lty = 3)
abline(a = -1.96, b = 0, lty = 3)

binnedplot(predict(poi.reg.ext), rstandard(poi.reg.ext))
dispersiontest(poi.reg.ext, trafo = 1)

yhat <- predict (poi.reg.ext, type = "response")
z <- (df$fupacts-yhat) / sqrt(yhat)
n = poi.reg.ext$df.null + 1
k = poi.reg.ext$df.null + 1 - poi.reg.ext$df.residual
cat("overdispersion ratio is ", sum(z^2) / (n-k), "\n")

cat("p-value of overdispersion test is ", pchisq(sum(z^2), n-k), "\n")
```

The estimated overdispersion is extremely high at 28.65, with over-dispersion ratio at  30.00404, and a p-value of over-dispersion test at  1.


### (c)

```{r}
df$c.bupacts <- (df$bupacts - mean(df$bupacts)) / (2 * sd(df$bupacts))
poi.reg.ext <- glm(fupacts ~ women_alone + sex + c.bupacts + couples + bs_hiv, family = quasipoisson, data = df)
display(poi.reg.ext)

ggplot(data=df, aes(x = couples, y = bupacts)) + geom_boxplot() + ylim(0, 50)

df$offset <- ifelse(df$bupacts == 0, 1, df$bupacts)
poi.reg.off <- glm(fupacts ~ women_alone + sex + couples + bs_hiv, offset = log(offset), family = quasipoisson, data = df)
display(poi.reg.off)
```

Singles tends to have unprotected sex more often than couples. We fit a Poisson model with the number of unprotected sex acts reported at the baseline as an offset. 

###(d)

Yes it should! Observations coming from the elements of couples is not i.i.d. THis yields an extremely high positive correlation between the answers of individuals that are a part of a couple.



# Problem 2

```{r}
require(arm)
require(foreign)
require(MASS)

df <- read.dta("/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/nes5200_processed_voters_realideo.dta")

df$partyid3 <- factor(df$partyid3, labels = c("democrats", "independents", "republicans", "apolitical"))
df$gender <- factor(df$gender, labels = c("male", "female"))
df$race <- factor(df$race, labels = c("white", "black", "asian", "native american", "hispanic", "other"))
df$south <- factor(df$south)
df$ideo <- factor(df$ideo, labels = c("liberal", "moderate", "conservative"))

x = df$partyid3
df <- df[!is.na(levels(x)[x]),]

df <- subset(df, partyid3 != "apolitical")
df$partyid3 <- factor(df$partyid3)
```


### (a)

```{r}
multi.log <- polr(partyid3 ~ ideo + race + age_10, Hess = TRUE, data = df)
summary(multi.log)
```


### (b)

```{r}
confint(multi.log)
```

age_10: For a one unit increase in age we expect a -0.11 increase in the expected value of partyid3.
ideo: moderates and especially conservatives are more likely to be republicans. In particular.
race: whites, and asianes are more likely to identify themselves as republicans, and blacks towards the democrat party.


### (c)

```{r}
residuals(multi.log)
```



# Problem 3

```{r}
require("arm")
require("foreign")
require("ggplot2")
require("VGAM")
require("gridExtra")

nsw <- read.dta("/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/NSW.dw.obs.dta")

nsw$sample <- factor(nsw$sample, labels = c("NSW", "CPS", "PSID"))
nsw$black <- factor(nsw$black)
nsw$hisp <- factor(nsw$hisp)
nsw$nodegree <- factor(nsw$nodegree)
nsw$married <- factor(nsw$married)
nsw$treat <- factor(nsw$treat)
nsw$educ_cat4 <- factor(nsw$educ_cat4, labels = c("less than high school", "high school", "sm college", "college"))

standardise <- function(X) {
    cols <- ncol(X)
    for (c in 1:cols) {
        if (is.numeric(X[, c])) {
            start <- ncol(X)
            c.c <- (X[, c] - mean(X[, c], na.rm=TRUE)) / (2 * sd(X[, c], na.rm = TRUE))
            X[start+1] <- c.c
            colnames(X)[start+1] <- paste0("c.", colnames(X)[c])
        }
    }
    return(X)
}

nsw <- standardise(nsw)
summary(nsw)

ggplot(data = nsw, aes(x = re78)) + geom_histogram(binwidth = (range(nsw$re78)[2] - range(nsw$re78)[1])/150)

modex <- function(x) {
    ux <- unique(x)
    ux[which.max(tabulate(match(x, ux)))]
}

print(paste0("The mode is: ", sprintf("$%3.2f", modex(nsw$re78))))
ggplot(nsw[nsw$re78 >= 25500 & nsw$re78 < 25700,], aes(x = re78)) + geom_histogram()

nsw$outcome <- rep(NA, nrow(nsw))
nsw$outcome <- ifelse(nsw$re78 >= 25564.669921875, 1, 0)
nsw$outcome <- factor(nsw$outcome, labels = c("lt", "gte"))

p1 <- ggplot(data=nsw, aes(x = outcome, y = age)) + geom_boxplot() + labs(title = "Outcome vs Age")
p2 <- ggplot(data=nsw, aes(x= outcome, y = re75)) + geom_boxplot() + labs(title = "Outcome vs re75")
grid.arrange(p1, p2, nrow = 1)

table(nsw$outcome, nsw$black)
table(nsw$outcome, nsw$hisp)
table(nsw$outcome, nsw$educ)

fit1 <- glm(outcome ~ c.age + c.educ + c.re75 + black + married, family = binomial(link = "logit"), data = nsw)
display(fit1)

predicted <- predict(fit1, nsw, type = "response")
y <- ifelse(nsw$re78 >= 25564.669921875, 1, 0)

error.rate <- mean((predicted > 0.5 & y == 0) | (predicted < .5 & y == 1))
print(paste0("Error rate: ", sprintf("%.2f%%", 100*error.rate)))

fit2.a <- vglm(re78 ~ c.age + c.educ + c.re75, tobit(Lower = 0, Upper = 25563), data = nsw, subset = re78 < 25564)
summary(fit2.a)

fit2.b <- vglm(re78 ~ c.age + c.educ + c.re75, tobit(Lower = 25564, Upper = Inf), data = nsw, subset = re78 >= 25564)
summary(fit2.b)

y.hat <- ifelse(predict(fit1, nsw) < 0.5, ifelse(predict(fit2.a, nsw) < 0, 0, predict(fit2.a, nsw)), predict(fit2.b, nsw))
y <- nsw$re78
print(paste0("RMSE: ", sprintf("%.2f", sqrt(mean((y - y.hat) ** 2)))))

ggplot(data = data.frame(cbind(y = y, y.hat = y.hat))) +
    geom_histogram(aes(x = y, fill = "y"), alpha = .35, binwidth = (range(y)[2] - range(y)[1])/150) +
    geom_histogram(aes(x = y.hat, fill = "y.hat"), alpha = .35, binwidth = (range(y)[2] - range(y)[1])/150)

y.hat <- ifelse(predict(fit1, nsw) < 0.5, ifelse(predict(fit2.a, nsw) < 0, 0, predict(fit2.a, nsw)), 25564.669921875) 
y <- nsw$re78
print(paste0("RMSE: ", sprintf("%.2f", sqrt(mean((y - y.hat) ** 2)))))

ggplot(data = data.frame(cbind(y = y, y.hat = y.hat))) +
    geom_histogram(aes(x = y, fill =" y"), alpha = .35, binwidth = (range(y)[2] - range(y)[1])/150) +
    geom_histogram(aes(x = y.hat, fill = "y.hat"), alpha = .35, binwidth = (range(y)[2] - range(y)[1])/150)
```

fit2.a: Underestimate the number of observations where earning in 1978 are zero.
fit2.b: Shifts the distribution towards values above $25,564.66.
Replaced the prediction of this model with the hard-coded value $25564.67, though further adjustments are needed to improve the fit of the model.



# Problem 4


### (a)

```{r}
df <- read.csv("/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/oscars.csv")
mcl <- polr(factor(Ch) ~ Nom + PrW + Gd + Gdr + DGA, Hess = TRUE, data = df)
summary(mcl)
```


### (b)

```{r}
Ch = as.numeric(df$Ch)
Nom = as.numeric(df$Nom)
PrW = as.numeric(df$PrW)
Gd = as.numeric(df$Gd)
Gdr = as.numeric(df$Gdr)
DGA = as.numeric(df$DGA)

data = as.matrix(c(Ch, Nom, PrW, Gd, Gdr, DGA))
mcl <- polr(factor(Ch) ~ Nom + PrW + Gd + Gdr + DGA, Hess = TRUE, data = data)
# plot(Ch, mcl) Error!
# hist(Ch, mcl) Error!
```
##### Not sure how to resolve this issue!

### (c)

```{r}
# hist(residuals(mcl)) Error!
# plot(residuals(mcl)) Error!
```
##### Not sure how to resolve this issue!


# Problem 5

```{r}
require(foreign)
require(nnet)
require(ggplot2)
require(reshape2)

ml <- read.dta("/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/hsbdemo.dta")
with(ml, table(ses, prog))
with(ml, do.call(rbind, tapply(write, prog, function(x) c(M = mean(x), SD = sd(x)))))
```

### (a)

```{r}
ml$prog2 <- relevel(ml$prog, ref = "academic")
test <- multinom(prog2 ~ ses + write, data = ml)
summary(test)

exp(coef(test))
head(pp <- fitted(test))
```
One-unit increase in the variable write is associated with the decrease in the log odds of being in general program vs. academic program in the amount of .058.
One-unit increase in the variable write is associated with the decrease in the log odds of being in vocation program vs. academic program. in the amount of .1136.
Odds of being in general program vs. in academic program will decrease by 1.163.
Odds of being in general program vs. in academic program will decrease by 0.533, although this coefficient is not significant.
Odds of being in vocation program vs. in academic program will decrease by 0.983.
Odds of being in vocation program vs. in academic program will increase by 0.291 although this coefficient is not significant.


### (b)

```{r}
dses <- data.frame(ses = c("low", "middle", "high"), write = mean(ml$write))
predict(test, newdata = dses, "probs")

dwrite <- data.frame(ses = rep(c("low", "middle", "high"), each = 41), write = rep(c(30:70), 3))
pp.write <- cbind(dwrite, predict(test, newdata = dwrite, type = "probs", se = TRUE))
by(pp.write[, 3:5], pp.write$ses, colMeans)
```


### (c)

```{r}
lpp <- melt(pp.write, id.vars = c("ses", "write"), value.name = "probability")
ggplot(lpp, aes(x = write, y = probability, colour = ses)) + geom_line() + facet_grid(variable ~ ., scales = "free")
```