---
title: '[STAT 4400] Exam-1'
author: "Michael Ghattas"
date: "2/22/2022"
output:
  
  word_document: default
  html_document: default
Refrences: https://cran.r-project.org/web/packages/mlogit/vignettes/e1mlogit.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1:



### (1)

True

### (2)

False

### (3)

True

### (4)

True

### (5)

False

### (6)

True

### (7)

True

### (8)

True

### (9)

True

### (10)

False

### (11)

True

### (12)

False

### (13)

False

### (14)

True

### (15)

False

### (16)

True

### (17)

True (IF by reflects we mean cause! As discussed with Bhawneet.)

### (18)

True

### (19)

False

### (20)

False (Correlated Explanatory Variables: If there are very many variables, it is likely that they will be highly correlated, meaning that some variables or sets of variables are measuring similar things. As discussed with Bhawneet.)




# Part 2:



### Problem - 1
```{r}
library(ggplot2)
head(msleep)
```
##### (1)

```{r}
df <- na.omit(msleep)
lmod = lm(sleep_total ~ brainwt, data = df)

ggplot(df, aes(brainwt, sleep_total, color = name)) +
  geom_point() + 
  geom_smooth(method = lm, color = "red") +
  theme_bw() + xlab("Average Brain Weight in KG") +   ylab("Total Amount of Sleep in Hrs") + 
  ggtitle("Sleep Times vs. Average Brain Weights")
```
##### (2)

```{r}
lmod = lm(sleep_total ~ awake + + bodywt + awake:bodywt + sleep_cycle, data = df)
summary(lmod)
plot(lmod)
```

##### (3)

```{r}
df$sleep_ratio <- (df$sleep_total / 24)
lmod = lm(sleep_total ~ sleep_ratio, data = df)
coef(lmod)
logitSR = (1 / (1 + exp(-1 * (-7.944109e-15 + (2.400e+01 * df$sleep_ratio))))); logitSR

lmod = lm(logitSR ~ log(brainwt), data = df)
summary(lmod)
```
Interpreting the results of our coefficients, we can see a negative correlation between the brain weight of the animal and the amount of sleep. For every additional KG of weight, there is a decrease of approximately 0.3% of sleep time needed per 24 hours.

```{r}
df$sleep_ratio <- (df$sleep_ratio * 24)
lmod = lm(sleep_total ~ sleep_ratio, data = df)
coef(lmod)
logitSR = (1 / (1 + exp(-1 * (3.177644e-15 + (1.000000e+00 * df$sleep_ratio))))); logitSR

lmod = lm(logitSR ~ log(brainwt), data = df)
summary(lmod)
```
We can see that returning from 24 hors to hours do not change our results and mantain the sleep time between 0 and 24, though never reaching 0 or 24.

##### (4)

```{r}
ggplot(df, aes(log(df$brainwt), df$sleep_ratio)) +
  geom_point() + 
  geom_smooth(method = lm, color = "red") +
  theme_bw() + xlab("Brain Weight in e^(KG)") +   ylab("Total Sleep in Hrs") + 
  ggtitle("Sleep Time per Day vs. Log(Brain Weight)")
```



### Problem - 2

```{r}
homeheat = read.csv('/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/Data/homeheat.csv')
head(homeheat)
```

##### (1)

```{r}
library("mlogit")
H <- dfidx(homeheat, choice = "depvar", varying = c(3:12))
m <- mlogit(depvar ~ ic + oc | 0, H)
summary(m)
```
Yes, the t-statistics are greater than 1.96, which is the critical level for 95% confidence level.

##### (2)

```{r}
coef(m)["oc"]/coef(m)["ic"]
```
The model implies that the decision-maker is willing to pay 73 cents in higher installation cost in order to reduce annual operating costs by $1.

##### (3)

```{r}
mc <- mlogit(depvar ~ ic + oc, H, reflevel = 'hp')
summary(mc)

apply(fitted(mc, outcome = FALSE), 2, mean)
```
Exact match: alternative-specific constants in a logit model insure that the average probabilities equal the observed shares.

##### (4)

```{r}
wtp <- coef(mc)["oc"] / coef(mc)["ic"]
wtp

r <- 1 / wtp
r
```
The willingness to pay is USD(4.56) for a $1 year stream of savings. The decision-maker applies a 22% discount rate, thus the results are certainly more reasonable than in the previous model.

##### (5)

```{r}
Hn <- H
Hn[idx(Hn, 2) == "hp", "ic"] <- 0.88 * Hn[idx(Hn, 2) == "hp", "ic"]
apply(predict(mc, newdata = Hn), 2, mean)
```
The share is predicted to rise to about 6.64% when rebates are given.

##### (6)

```{r}
plot1 = ggplot(homeheat, aes(ic.gc + ic.gr + ic.ec + ic.er + ic.hp, oc.gc  + oc.gr  + oc.ec  + oc.er + oc.hp, color = idcase)) +
  geom_point() + 
  geom_smooth(method = lm, color = "red") +
  theme_bw() + xlab("Installation Cost") +   ylab("Annual Operating Cost") + 
  ggtitle("Installation vs. Operational Cost")

plot2 = ggplot(homeheat, aes(agehed, income)) +
  geom_point(shape = 21, color = "darkgoldenrod4", fill = "darkgoldenrod3", size = 5) + 
  theme_light() + xlab("Age") +   ylab("Income Class") + 
  ggtitle("House-Head Age vs. Income Class")

library(gridExtra)
grid.arrange(plot1, plot2, ncol = 2)
```
