---
title: '[STAT 4400] HW-1'
author: "Michael Ghattas"
date: "1/17/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Question-1

$n = 3$

$\bar{y} = \frac{1}{n}\sum_{i = 1}^{n}y_i = \frac{1 + 2 + 3}{3} = \frac{6}{3} = 2$ 

$\bar{x} = \frac{1}{n}\sum_{i = 1}^{n}x_i = \frac{2 + 0 + 4}{3} = \frac{6}{3} = 2$


### (a)

$\hat{\beta_0} = \bar{y} - \hat{\beta_1}\bar{x}$

$\hat{\beta_1} = \frac{\sum_{i = 1}^{n}(y_i - \bar{y})(x_i - \bar{x})}{\sum_{i = 1}^{n}(x_i - \bar{x})^2}$

$\hat{\sigma^2} = \frac{1}{n - 2}\sum_{i = 1}^{n}(y_i - \hat{\beta_0} - \hat{\beta_1}x_i)^2$


### (b)

$\hat{\beta_1} = \frac{\sum_{i = 1}^{3}(y_i - 2)(x_i - 2)}{\sum_{i = 1}^{3}(x_i - 2)^2} = \frac{(-1 + 0 + 1)(0 - 2 + 2)}{(0)^2 + (-2)^2 + (2)^2} = \frac{(0)(0)}{4 + 4} = \frac{0}{8} = 0$

$\hat{\beta_0} = 2 - (0)\bar{x} = 2 - 0 = 2$

$\hat{\sigma^2} = \frac{1}{3 - 2}\sum_{i = 1}^{3}(y_i - 2 - (0)x_i)^2 = \frac{1}{1}\sum_{i = 1}^{3}(y_i - 2)^2 = 1[(-1)^2 + (0)^2 + (1)^2] = 1 + 0 + 1 = 2$


### (c)

$\hat{\epsilon_i} = y_i - \hat{y_i} = y_i - \hat{\beta_0} - \hat{\beta_1}x_i = y_i - 2 - 0 = y_i - 2$
$SSE = \sum_{i = 1}^{3}\hat{\epsilon_i}^2 = (1 - 2)^2 + (2 - 2)^2 + (3 - 2)^2 = (-1)^2 + (0)^2 + (1)^2 = 1 + 0 + 1 = 2$

$\hat{y_i} = \hat{\beta_0} - \hat{\beta_1}x_i + \hat{\epsilon_i} = 2 - 0 + \hat{\epsilon_i} = 2 + \hat{\epsilon_i}$
$SSR = \sum_{i = 1}^{3}(\hat{y_i} - 2)^2 = ((2 + (1 - 2)) - 2)^2 + ((2 + (2 - 2)) - 2)^2 + ((2 + (3 - 2)) - 2)^2 = ((2 - 1) - 2)^2 + ((2 - 0) - 2)^2 + ((2 + 1) - 2)^2 = (1 - 2)^2 + (2 - 2)^2 + (3 - 2)^2 = (-1)^2 + (0)^2 + (1)^2 = 1 + 0 + 1 = 2$

$SST = \sum_{i = 1}^{3}(y_i - 2)^2 = (1 - 2)^2 + (2 - 2)^2 + (3 - 2)^2 = (-1)^2 + (0)^2 + (1)^2 = 1 + 0 + 1 = 2$

$R^2 = 1 - \frac{SSE}{SST} = \frac{2}{2} = 1$


### (d)

$H_0: \beta_1 = 0$

$H_1: \beta_1 \neq 0$

$\beta_1* = 0$

$\hat{\beta_1}$ ~ N(0, $\tau$)

$\hat{\tau} = \sqrt{\frac{\sigma^2}{\sum_{i = 1}^{3}(x_i - \bar{x})^2}} = \sqrt{\frac{2}{\sum_{i = 1}^{3}(x_i - 2)^2}} = \sqrt{\frac{2}{(1 - 2)^2 + (2 - 2)^2 + (3 - 2)^2}} = \sqrt{\frac{2}{(-1)^2 + (0)^2 + (1)^2}} = \sqrt{\frac{2}{1 + 0 + 1}} = \sqrt{\frac{2}{2}} = \sqrt{1} = 1$

$\hat{t} = \frac{\hat{\beta_1}}{\hat{\tau}}$

$p[|t| \geq t*] = \alpha$

```{r}
y = c(1 , 2 , 3)
x = c(2 , 0 , 4)
alpha = 0.5
n = length(x)

beta.0.hat = 2
beta.1.hat = 0
sigma2.hat = 2
std.t = dt(x , (n - 2))
t.star = qt((1 - alpha / 2) , df = (n - 2))

y.hat = beta.0.hat + (beta.1.hat * x)
eps.hat = y - y.hat
SSE = sum(eps.hat ** 2)
sigma2.hat = SSE / (n - 2)
tau.hat = sqrt(sigma2.hat / sum((x - mean(x)) ** 2))
t.hat = beta.1.hat / tau.hat

t.star
t.hat

print("(|t| < t*) Thus we fail to reject the the null-hypothesis")
```


### (e)

```{r}
lmod = lm(y ~ x)
plot(x, y)
abline(lmod, col = "red")
```



# Question-2

$0 \leq x \leq 50$
$\bar{x} = 35$
$\sigma_x = 10$

$y = a + bx$
$\bar{y} = a + b\bar{x} = a + 35b$

$\sigma_y = 15$
$\bar{y} = 100$

$\sigma_y = |b|\sigma_x = 10|b|$
$15 = 10|b| \to |b| = \frac{15}{10} = 1.5 \to b = \pm1.5$

$a + 35b = 100 \to a + 35(\pm1.5) = 100$
$a + 35(-1.5) = 100 \to a - 52.5 = 100 \to a = 100 + 52.5 = 152.5$
$a + 35(1.5) = 100 \to a + 52.5 = 100 \to a = 100 - 52.5 = 47.5$


### (a)

$y = 152.5 - 1.5x$


### (b)

$y = 47.5 + 1.5x$


### (c)

##### I would recommend the solution from part (a), as it has an increasing function that exhibits a positive correlation.



# Question-3


### (a)

```{r}
set.seed(123)

var1 = rnorm(1000, mean = 0, sd = 1)
var2 = rnorm(1000, mean = 0, sd = 1)

lmod1 = lm(var1 ~ var2)
summary(lmod1)

print("Yes, based on the p-value generated we can conclude that the slope coefficient var2 is statistically significant.")
```


### (b)

```{r}
set.seed(321)

z.scores <- rep (NA, 100)

for (k in 1:100) 
{
  var1 <- rnorm (1000 ,0 ,1) 
  var2 <- rnorm (1000 ,0 ,1)
  fit <- lm (var2  ~ var1)
  z.scores[k] <- coef(fit )[2] / summary(fit)$coef[2,"Std. Error"]
}

alpha = .05
cutoffn = qnorm((1 - alpha) / 2, lower.tail=TRUE)
sum(abs(z.scores) > cutoffn)

for (k in 1:100) 
{
  result = sum(abs(z.scores) < 1.96)
}

result
print("95 estimated slope coefficients are statistically significant at the α = .05 level of significance.")
```



# Question-4


### (a)

```{r}
library(haven)
data <- read_dta("/Users/Home/Documents/Michael_Ghattas/School/CU_Boulder/2022/Spring 2022/STAT - 4400/HW/1/child.iq.dta")
head(data)

lmod = lm(ppvt ~ momage, data = data)
summary(lmod)

library(ggplot2)
ggplot(data, aes(momage, ppvt)) +
geom_point(shape = 21, color="steelblue4", fill="steelblue3", size = 3,
alpha=0.5,show.legend = FALSE) +
theme_light() + xlab("Age (Mother)") + ylab("Child (Test Scores)") +
geom_smooth(method = lm, color="darkred", se=FALSE)

print("Based on our summary and plots, it seems that the mother's age is a significant predictor, though not enough on its own to find a direct correlation. That being said, the plot clearly shows that the majority of the observations with a higher test score belong to the mothers in their late 20's. Thus mothers should give birth in their late 20s.")
```


### (b)

```{r}
lmod = lm(ppvt ~ momage + educ_cat, data = data)
summary(lmod)

library(ggplot2)
ggplot(data, aes(educ_cat, ppvt)) +
geom_point(shape = 21, color="steelblue4", fill="steelblue3", size = 3,
alpha=0.5,show.legend = FALSE) +
theme_light() + xlab("Education (Mother)") + ylab("Child (Test Scores)") +
geom_smooth(method = lm, color="darkred", se=FALSE)

print("Based on our summary and plots, it seems that the mother's education is a strong and significant predictor. That being said, the plot clearly shows that the majority of the observations with a higher test score belong to the mothers having completed a high-school education.")
```


### (c)

```{r}
data$mom.hs <- ifelse(data$educ_cat >= 2, 1, 0)

lmod <- lm(ppvt ~ (mom.hs * momage), data = data)
summary(lmod)


spread <- ifelse(data$mom.hs == 1, "red", "black")
plot(data$momage, data$ppvt, xlab = "Age (Mother)", ylab = "Child (Test Score)", col = spread, pch = 1)
curve(cbind(1, 1, x, 1 * x) %*% coef(lmod), add = TRUE, col = "orange")                                   # Mother finished hs
curve(cbind(1, 0, x, 0 * x) %*% coef(lmod), add = TRUE, col = "grey")                                     # Mother did not finish hs
```



### (d)

```{r}
lmod <- lm(ppvt ~ momage + educ_cat, data = data[1:200, ])
summary(lmod)

pmod <- predict(lmod, data[201:400, ])
plot(pmod, data$ppvt[201:400], xlim = c(70, 110), xlab = "Predicted", 
    ylab = "Actual", col = "blue")
abline(a = 0, b = 1, col = "red")
```



# Question-5


### (a)

```{r}
library(faraway)
data(prostate)
head(prostate)

lmod = lm(log(lpsa) ~ log(lcavol) + ., data = prostate)
summary(lmod)

confint(lmod, 'log(lcavol)', level = 0.95)
```


### (b)
```{r}
lmod = lm(log(lpsa) ~ log(lcavol) + lcavol + lbph + svi, data = prostate)
summary(lmod)

confint(lmod, 'log(lcavol)', level = 0.95)
```


### (c)

##### The model from part (a) has a slightly better fit, as the R^2 value is slightly higher, indicating a better fitted model.